
# Lab1 Answers File
--
Please write your answers into the space below the questions. Knit this file when you are done to check that all code runs. Remember to rename the file to contain *your actual* Last and First name and **submit both the Rmd and Html file to canvas**.
--
#### Amatullah Mir

**1.1**
```{r}
(pi)^8
```

**1.2**
```{r}
(4*10^9)*(pi^3)
```

**1.3**
<span style="color:blue"> It outputs the following error: "Error: attempt to apply non-function." This is because the function is missing the asterisk before the parentheses. It should be written 2*(5+7). </span>

**2.1**
```{r}
charvect <- c("Amatullah", "Mir", "BIOS20172")
```

**3.1**
```{r eval = FALSE}
myVec[1:8]       # 10 20 30 40 50 60 70 80. This returns the vector's elements 1 through 8.
myVec[-2]        # 10 30 40 50 60 70 80 90 100. This omits the second element of the vector.
myVec[c(2,4,6)]  # 20 40 60. This accesses the 2nd, 4th, and 6th elements.
myVec[2] <- 15   # 10  15  30  40  50  60  70  80  90 100. This inserts 15 into the vector.
4+myVec          # 14  19  34  44  54  64  74  84  94 104. This adds 4 to each value in the vector.
```

**3.2**
```{r}
N_0 <- 100
t <- 3

(N_0)*2^(t)
            
```

**3.3**
```{r}
ejectionVolume <- 70
HeartRate <- c(seq(60, 100, by = 5))
200*ejectionVolume*HeartRate 
```

**4.1**
<span style="color:blue"> The difference between the two matrices are that they are filled differently. The byrow argument specifies that the matrix must be filled in a certain way: by rows instead of filling by columns first. </span>

**4.2**
```{r}
x <- matrix(nrow = 5, ncol = 5, data = 1:25)
mean(x) # 0.8236657 -> calculates by using all the elements in the matrix.
sd(x) # 7.359801
sum(x) # 325
```
<span style="color:blue"> R performs these calculations by using all of the elements in the matrix. </span>

**4.3**
```{r}
amatullah <- matrix(nrow = 3, ncol = 4, data = 1:12)
amatullah
dim(amatullah)
```
<span style="color:blue"> The dim() function returns two numbers: the first represents the number of rows present in the matrix, and the second represents the number of columns present in the matrix. </span>

**4.4**
```{r}
mydat <- c(1, 2, 3, 5, 7, 11, 13, 17, 19)
myMatrix <- matrix(nrow = 3, ncol = 3, data = mydat)
myMatrix # standard matrix of 9 prime numbers
myMatrix[2,]<- c(4, 6, 8)
myMatrix # second row is replaced by the first three composite numbers.
```

**5.1**
<span style="color:blue"> < compares values (less than), > also compares (greater than). <= compares values according to whether they are less than or equal to, and >= compares values according to whether they are greater than or equal to. != determines whether or not two values are not equal to each other. </span>

**5.2**
```{r}
dresser <- 1
room <- 1
closet <- 2

dresser == room
room == closet

(closet > room) && (room <= dresser)
(closet < room) && (room <= dresser)

(closet > room) || (room <= dresser)
(closet < room) || (room > dresser)
```

**6.1**
<span style="color:blue"> FALSE == 0 returns FALSE, and TRUE == 1 returns TRUE. </span>

**6.2**
```{r}
temp <- 90 # arbitrary value for temperature
threshold <- 102
if(temp >= threshold){
  print("Seek medical attention")
} else {
  print("Do nothing")
}
```

**6.3**
<span style="color:blue"> The string "B" is outputted, since the value that R might've interpreted as the input was not 0. </span>

**6.4**
```{r}
temp <- 103 # arbitrary value for temperature
threshold1 <- 99
threshold2 <- 102
if(temp >= threshold1){
  if(temp >= threshold2){
  print("Seek medical attention ASAP!")
} else {
  print("Seek medical attention.")
}
} else {
  print("Do nothing")
}
```

**7.1**
<span style="color:blue"> When x is even, x%%y is going to be 0, and when it is odd, it will be 1. </span>

**7.2**
```{r}
for(i in 1:20){
  if(i%%2==1 || i==12){
  print(i)
  }
}
```

**7.3**
```{r}
for(i in letters){
  if(i=="c"){
    print("The third letter of the alphabet.")
  } else {
    print("Not the third letter of the alphabet.")
  }
}
```
<span style="color:blue"> for loops can also iterate over letters, not just numbers. </span>

**7.4**
<span style="color:blue"> The code is going to yield an infinite loop, since there is no condition to stop it from continually running. </span>

```{r eval = FALSE}
x <- 1
while(x <= 5){
  print(x + x)
  x <- x + 2
}
```

############################## Review ##############################

1.
```{r}
dengue <- read.fasta(file = 'dengue_seq.txt')
dengue_vec <- as.vector(dengue[[1]])
dengue_vec[1:20]
```

2.
```{r}
length3 <- 0
for(i in dengue_vec){
  length3 = length3+1
  }
length3
```

3.
```{r}
length2 <- 0
for(i in dengue_vec){
  if(i=="c"){
    length2 = length2+1
  }
}
length2
```

4.
```{r}
length1 <- 0
for(i in dengue_vec){
  if(i=="g"){
    length1 = length1+1
  }
}
length1
```

5.
```{r}
lengths <- (length1 + length2)
GC_content = (lengths/length3)*100
GC_content
```

6.
```{r}
abs(-10) # outputs the absolute value of a number
sqrt(25) # outputs the square root of a number
ceiling(100.98) # outputs the rounded up version of the integer inputted.
floor(6.78) # outputs the rounded down version of the integer inputted. 
trunc(6.953452) # eliminates everything to the right of the decimal.
```
<span style="color:blue"> The absolute value, square root, ceiling, floor, and truncate function. abs takes the absolute value of a number. Sqrt takes the square root of an inputted number. Ceiling rounds the number up, floor rounds the number down, and truncate eliminates the decimals of the number. </span>

7. <span style="color:blue"> This has to do with booleans in R and their integration of integers. 1 is always true and all other numbers, including 0, 2, 3, and so on. "if(2){print("A")}else{print("B")}" is true, because no number is inputted, the argument is false. 2 and that argument will both be FALSE, which is why "A" is returned. </span>

